---
title: "MATH410HW7"
author: "Tyler Millan"
date: "5/29/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Desktop/MATH410/HW7")
library(readxl)

#The below function was borrowed from NICOLE RADZIWILL from https://qualityandinnovation.com/2015/03/16/one-proportion-z-test-in-r/

z.test <- function(x,n,p=NULL,conf.level=0.95,alternative="less") {
   ts.z <- NULL
   cint <- NULL
   p.val <- NULL
      phat <- x/n
      qhat <- 1 - phat
      # If you have p0 from the population or H0, use it.
      # Otherwise, use phat and qhat to find SE.phat:
      if(length(p) > 0) { 
         q <- 1-p
         SE.phat <- sqrt((p*q)/n) 
         ts.z <- (phat - p)/SE.phat
         p.val <- pnorm(ts.z)
         if(alternative=="two.sided") {
            p.val <- p.val * 2
         }
         if(alternative=="greater") {
            p.val <- 1 - p.val
         }
      } else {
      # If all you have is your sample, use phat to find
      # SE.phat, and don't run the hypothesis test:
         SE.phat <- sqrt((phat*qhat)/n)
      }
      cint <- phat + c( 
         -1*((qnorm(((1 - conf.level)/2) + conf.level))*SE.phat),
         ((qnorm(((1 - conf.level)/2) + conf.level))*SE.phat) )
   return(list(estimate=phat,ts.z=ts.z,p.val=p.val,cint=cint))
}
```

8.28, 8.37, 8.39, 8.71, 8.92

8.28
```{r 8.28}

x=62380
n=159949
p_hat<-x/n
se<- sqrt(p_hat*(1-p_hat)/n)
zcrit<-2.576
moferror<-se*zcrit

```

a) The margin of error for 99% confidence is `r moferror`.

b) This margin of error tells us possible errors in our study because the above accounted for just the random sampling error, while this accounts for anything outside of this.

8.37
```{r 8.37}
x=35
n=50
pcoffee = x/n

coffeetest2<-z.test(x=35, n = 50,p=.5,alternative = "greater")

coffeetest2$ts.z
```
a)
We are looking to test the claim that people prefer the taste of fresh brewed coffee over the taste of instant coffee.
The null hypothesis to test this will be that the proportion is equal to .5, and our alternative will be that the proportion is greater than .5.
H0:p=50
Ha:p>50
After running the one proportion test on this we get a pvalue of `r coffeetest2$p.val`. We also get a large-sample z statistic of `r coffeetest2$ts.z`.

b)

```{r 8.37b}
cord.x <- c(2.83,seq(2.83,3,0.01),3) 
cord.y <- c(0,dnorm(seq(2.83,3,0.01)),0) 
 

curve(dnorm(x,0,1), xlim=c(-3,3), main='Standard Normal') 
 

polygon(cord.x,cord.y,col='skyblue')


```

The above graph's shaded area starts at 2.83 with a p-value of `r coffeetest2$p.val`.

c)After running the one proportion test on this we get a pvalue of `r coffeetest2$p.val` so we reject at the 5% level that people are indifferent between instant and fresh brewed coffee, implying it is significant at the 5% level. Given this a practical conclusion we can come to is that people prefer fresh brewed coffee to instant coffee.

8.39
```{r 8.39}

cointest<-z.test(x=5067, n = 10000,p=.5,alternative = "two.sided")

cord.x <- c(-3,seq(-3,-1.34,0.01),-1.34) 
cord.y <- c(0,dnorm(seq(-3,-1.34,0.01)),0)
 
cord1.x<- c(1.34,seq(1.34,3,0.01),3) 
cord1.y <- c(0,dnorm(seq(1.34,3,0.01)),0)
curve(dnorm(x,0,1), xlim=c(-3,3), main='Standard Normal') 
 

polygon(cord.x,cord.y,col='skyblue')
polygon(cord1.x,cord1.y,col='skyblue')
```
a) To test whether Kerrich's coin comes up heads is not .5 we test using a one proportion test where our null hypothesis is that the proportion is equal to .5 (H0:p=.5) and our alternative is that it is not equal to .5(Ha:p!=.5). After running our test we get a z score of `r cointest$ts.z` and a pvalue of .1802 which means it is not significant at the 5% level and we fail to reject our null hypothesis that the proportion is equal to .5. 

b) The range for our 95% confidence interval that the probabilities of heads that wont be rejected at the 5% level falls between (`r cointest$cint`)

8.71
```{r 8.71}
res <- prop.test(x = c(48, 52), n = c(60, 132),correct = FALSE, conf.level = 0.90)
sefemale<- sqrt((res$estimate[1]*(1-res$estimate[1]))/60)
semale<- sqrt((res$estimate[2]*(1-res$estimate[2]))/132)

```

a) The proportion of juvenile references for females is `r res$estimate[1]`, while the proporion of juvenile references for males is `r res$estimate[2]`. The standard error of juvenile references for females is `r sefemale`, while the standard error of juvenile references for males is `r semale`.
8.92.

b)The range for our 90% confidence interval that the proportion of juvenile references at the 10% level falls between (`r res$conf.int`).

c) Sincce we want to test the significance of the two being equal we will use a two proportion z test where the null hypothesis will be that the two proportions are equal(H0: p1=p2) and where the alternative hypothesis is that the two proportions are not equal(Ha: p1!=p2). After running the test we are given a z score of 5.22 and a pvalue of `r res$p.value` which is less than the significance level of .1 so we reject the null hypothesis, which means it is sigificanct at the 10% level, meaning we can conclude there is a difference in the proportion of female references that are girls and male references that are boys.


```{r 8.92}

genderdata <- read_excel("ex08-92genderb.xls")
#1
prop.test(x = c(2, 7), n = c(5, 35),alternative = "less")
#2
prop.test(x = c(5, 18), n = c(7, 63),alternative = "less")
#3
prop.test(x = c(25, 14), n = c(56, 65),alternative = "less")
#4
prop.test(x = c(11, 19), n = c(76, 157),alternative = "less")
#5
prop.test(x = c(2, 12), n = c(3, 43),alternative = "less")
#6
prop.test(x = c(48, 52), n = c(60, 132),alternative = "less")
#7
prop.test(x = c(38, 70), n = c(40, 72),alternative = "less")
#8
prop.test(x = c(5, 6), n = c(18, 33),alternative = "less")
#9
prop.test(x = c(48, 128), n = c(48+24, 128+48),alternative = "less")
#10
prop.test(x = c(13, 32), n = c(13+5, 32+95),alternative = "less")
```

**answer:** We fail to reject reject the null hypothesis for all textbook numbers that there is a difference in proportion between juvenile references. We can conclude that there is no pattern that suggests the gender of an other being asccociated with the results of datat collected.  